---
title: "p8105_hw2_kq2127"
author: "Kristal Quispe"
date: "9/29/2019"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

## Problem 1

Read and clean Mr. Trash Wheel sheet

```{r}
mtw_data = 
  read_excel(
    path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range ="A2:N406") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))
 
mtw_data

    
```

Read, clean and combine 2017-2018 Precipitation data

```{r}
precip17_dt = 
  read_excel(
    path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2017)

precip17_dt



precip18_dt = 
  read_excel(
    path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2018)

precip18_dt



precip_dt =
  bind_rows(
    precip17_dt,precip18_dt) %>%
  janitor::clean_names() %>% 
  mutate(month = month.name[month])

precip_dt
```

The mtw_data data set has 344 observations. The most important variables in this data set are dumpster, year, weight_tons, volume_cubic_yards, and sports_balls. The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(mtw_data,year == 2017), sports_balls))`. The precip_dt data set has 24 observations. The most important variables in this data set are year, month and total.  The total precipitation in 2018 is `r sum(pull(precip18_dt, total))` units. 

## Problem 2

```{r}
pm_dt = read_csv(file="./data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year)) %>% 
  select(year, month, prez_gop, prez_dem, everything(), -day) %>% 
  pivot_longer(
    prez_gop:prez_dem,
    names_to = "president",
    names_prefix = "prez_",
    values_to = "value") %>% 
  filter(value == 1) %>% 
  select(year, month, president, everything(), -value)

pm_dt

snp_dt = read_csv(file="./data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year)) %>% 
  select(year, month, everything(), -day)

snp_dt

une_dt = read_csv(file="./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment") %>% 
  mutate(
    month = recode(
      month,
      "jan" = "January",
      "feb" = "February",
      "mar" = "March",
      "apr" = "April",
      "may" = "May",
      "jun" = "June",
      "jul" = "July",
      "aug" = "August",
      "sep" = "September",
      "oct" = "October",
      "nov" = "November",
      "dec" = "December"))

une_dt

fj_dt = left_join(pm_dt, snp_dt, by = c("year", "month"))

pb2_dt = left_join(fj_dt, une_dt, by = c("year", "month"))

pb2_dt
```

   The data in the pols-month csv file was cleaned and tidyed to become the pm_dt data set which has 817 observations and 9 variables. This data set contains political information regarding how many house, senate and governor seats were then being occupied by Democrats vs. Republicans for a given year and month. The data ranges from year 1947 to 2015. The pm_data set also has a column called president that shows us the party the president belonged to for every month of a year. The key variables in this data set are year, month and president. 
   The snp csv file was cleaned and tidyed to become the snp_dt data set which contains 787 observation and 3 variables. The data ranges from year 1950 to 2015. This data set contains information about the close for a given year and month. The key variables in this data set are year, month and close. 
   The unemployment data was cleaned and tidyed and became the une_dt data set, which has 816 observations and 3 variables. This data describes the unemployent rate for every month of the year between 1948 to 2015. The key variables in this data set are year, month and unemployment. 
   All three of these data sets were merged together on year and month to create one data set called pb2_dt. This data set has 817 observations and 11 variables. The time range for this data set is from 1947 to 2015, and the key variables are month, year, president, close, and unemployment. 


## Problem 3

```{r}
babynames_dt = read_csv(file="./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    childs_first_name = tolower(childs_first_name),
    ethnicity = recode(ethnicity,
      "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
      "BLACK NON HISP" = "BLACK NON HISPANIC",
      "WHITE NON HISP" = "WHITE NON HISPANIC")) %>% 
  distinct()

babynames_dt     

olivia_dt =
  filter(babynames_dt, childs_first_name == "olivia") %>% 
  select (ethnicity, year_of_birth, rank) %>% 
  pivot_wider(names_from = year_of_birth, values_from = rank)

olivia_dt

boy_dt =
  filter(babynames_dt, gender == "MALE", rank == 1) %>% 
  select (ethnicity, year_of_birth, childs_first_name) %>% 
  pivot_wider(names_from = year_of_birth, values_from = childs_first_name)
  
boy_dt

sp_dt = 
  filter(
    babynames_dt, 
    gender == "MALE", 
    ethnicity == "WHITE NON HISPANIC", 
    year_of_birth == 2016 ) %>% 
  ggplot(aes(x = rank, y = count)) + geom_point()

sp_dt

```


